---
title: "Case study: Cyclistic How to tranform the casual riders to manual riders"
output: html_notebook
editor_options:
  markdown: 
    wrap: sentence
---

## Part 1: Ask

1.  Business task statement As a junior data analyst at Cyclistic, my task is to analyze how annual members and casual riders use Cyclistic bikes differently.
    The goal is to uncover behavioral patterns and trends in ride length, frequency, day-of-week usage, and other relevant variables.
    These insights will support the marketing team in designing targeted strategies to convert more casual riders into annual members and thus increase long-term revenue for the company.

2.  My stakeholders:

-   Lily Moreno - Director of Marketing – needs insights to guide the new marketing campaign
-   Marketing Analytics Team - Will use your analysis to measure performance and impact of the campaign
-   Cyclistic Executive Team - Final decision-makers who must approve any proposed strategy

3.  The business context There are over 5800 bikes have been shared and over 692 docking stations.
    The users are devided two type: Annual members and casual riders (single/day-pass users).
    Depening on their purposes, but the large scale in their users uses bike for their daily traveling to work.
    There are 4 types of bike: Standard bikes, e-bikes, hand tricycles, cargo bikes (8% use assistive options)

4.  The business questions

-   "How do annual members and casual riders use Cyclistic bikes differently?"
-   Why would casual riders choose to become members?
-   How can digital media be used to influence their decision?

## Part 2: Prepare and process dataset

1.  Loading dataset The dataset is located in "Cloud/project/dataset" on the Positcloud website. Let's see what the data is organized:

```{r setting packages}
install.packages('tidyverse')
library(tidyverse)
install.packages('readr') #when install tidyverse was failed, cannot install completely readr
library(readr)

# inport data
divvy_2019_q1 <- read_csv('dataset/Divvy_2019_Q1.csv')
divvy_2020_q1 <- read_csv('dataset/Divvy_2020_Q1.csv')
```

2.  Exploring the dataset The dataset is ROCCC:

-   Reliable: cannot judge because it is hypothetical data
-   Original: yes.
-   Comprehensive: yes
-   Current: yes, the dataset was setting for the questions in the 2020 context.
-   Cited: yes. Because these are the datasets was provided in my Course, so I don't have any problems with the authority or privacy. And I also recheck them. The data have any error or how about the dataset's integrity?

```{r}
install.packages('skimr')
library(skimr)
divvy_2019_q1 %>% 
skim_without_charts()
```

Dataset of quater 1 in 2019: 12 columns and 365069 rows =\> the gender has 19711 missing values and birthyear has 18023 missing value.
So, I decide to assign "Unknown" for the missing value.

```{r}
divvy_2019_q1$gender <- ifelse(is.na(divvy_2019_q1$gender), "Unknown", divvy_2019_q1$gender)
divvy_2019_q1$birthyear <- ifelse(is.na(divvy_2019_q1$birthyear), "Unknown", divvy_2019_q1$birthyear)
```

```{r}
table(divvy_2019_q1$usertype)
prop.table(table(divvy_2019_q1$usertype))

```

ALmost the data in divvy_2019_q1 is about subcriber, if what I think is true, this dataset is almost about the manual user and just have small part user have not yet subcribed membership (6.34%).
So, in the some sides, the rate tranformation users (new user to subriber) is high.

```{r}
table(divvy_2019_q1$gender)
prop.table(table(divvy_2019_q1$gender))
```

What wrong with female, their husband or boyfriend ride the bike?
Or they prefer other type vehicles?

```{r}
divvy_2020_q1 %>% 
  skim_without_charts()
```

Divvy_2020_q1: in end_station_name, end_station_id, end_lat, end_lng columns, each of them has one missing value.
I assign "Unknown" for missing value.

```{r}
divvy_2020_q1$end_station_name <- ifelse(is.na(divvy_2020_q1$end_station_name), "Unknown", divvy_2020_q1$end_station_name)
divvy_2020_q1$end_station_id <- ifelse(is.na(divvy_2020_q1$end_station_id), "Unknown", divvy_2020_q1$end_station_id)
divvy_2020_q1$end_lat <- ifelse(is.na(divvy_2020_q1$end_lat), "Unknown", divvy_2020_q1$end_lat)
divvy_2020_q1$end_lng <- ifelse(is.na(divvy_2020_q1$end_lng), "Unknown", divvy_2020_q1$end_lng)
```

The rate of member type and gender

```{r}
table(divvy_2020_q1$member_casual)
prop.table(table(divvy_2020_q1$member_casual))
```

I have a hypothety that the firm change the way deviding customer: casual == customer, subcriber == member.

```{r}
table(divvy_2019_q1$gender)
prop.table(table(divvy_2019_q1$gender))
```

3.  Creating new columns

-   What time for each trip: ride_length: the duration of each ride in minutes

```{r}
library(dplyr)
divvy_2019_q1 <- divvy_2019_q1 %>%
  mutate(ride_length = as.numeric(difftime(end_time, start_time, units = "mins")))
divvy_2020_q1 <- divvy_2020_q1 %>%
  mutate(ride_length = as.numeric(difftime(ended_at, started_at, units = "mins")))


```

Remove if the ride length \<= 0:

```{r}
divvy_2019_q1 <- divvy_2019_q1 %>% 
  filter(ride_length > 0)
divvy_2020_q1 <- divvy_2020_q1 %>% 
  filter(ride_length >0)
```

What day on the week the trip was started?:
day_of_week: the weekday when the ride started

```{r}
divvy_2019_q1 <- divvy_2019_q1 %>% 
  mutate(day_of_week = weekdays(as.Date(start_time)))
divvy_2020_q1 <- divvy_2020_q1 %>% 
  mutate(day_of_week = weekdays(as.Date(started_at)))
```

4.  Analyzing dataset

-   Calculating the statistic of each member type. **divvy_2019_q1**

```{r}
divvy_2019_q1 %>% 
  group_by(usertype) %>% 
  summarise(
    mean_ride = mean(ride_length, na.rm = T),
    median_ride = median(ride_length, na.rm = T),
    max_ride = max(ride_length, na.rm =T),
    min_ride = min(ride_length, na.rm = T)
  )
```

**divvy_2020_q1**

```{r}
divvy_2020_q1 %>% 
  group_by(member_casual) %>% 
  summarise(
    mean_ride = mean(ride_length, na.rm = T),
    median_ride = median(ride_length, na.rm = T),
    max_ride = max(ride_length, na.rm =T),
    min_ride = min(ride_length, na.rm = T)
  )
```

We can see that in the both periods (q1 2019, 2020), the casual users often have longer distances than member users.

###Calculating the number of trips for each weekday.
**divvy_2019_q1**

```{r}
trip_2019 <- divvy_2019_q1 %>% 
  group_by(day_of_week, usertype) %>% 
  summarise(num_trip_of_weekday = n(),
            mean_time_trip = mean(ride_length, na.rm = T)) %>% 
  arrange(day_of_week, usertype)
trip_2019
```

**divvy_2020_q1**

```{r}
trip_2020 <- divvy_2020_q1 %>% 
  group_by(day_of_week, member_casual) %>% 
  summarise(num_trip_of_weekday = n(),
            mean_time_trip = mean(ride_length, na.rm = T)) %>% 
  arrange(day_of_week, member_casual)
trip_2020
```

The number of member user'trips focus on weekday, around 5000 to 6000 units, an the number of casual user's trips skyrocket on the weekend =\> these are evidences for the hythority: member users are almost worker and they use vehicle for daily traveling, the purpose of casual users is for entertainment.
the time in each trip in both dataset provide me that the member user usualy take the short trip, casual travel long time.

## Part 4: Share - Creating Viz

Viz show the relationships between weekdays and the number of trips in 2019

```{r}
library(ggplot2)
ggplot(trip_2019, aes(x = day_of_week, y = num_trip_of_weekday, fill = usertype)) +
  geom_col() + 
  labs(title = "The number of weekday 2019")
```

The Viz show the relationship between weekdays and the trip time

```{r}
library(ggplot2)
ggplot(trip_2019, aes(x = day_of_week, y = mean_time_trip, fill = usertype)) +
  geom_col() + 
  labs(title = "The trip time of weekday 2019")
```

We can see that the trips during the week are all by the subscriber group, they are the majority, only on weekends do they use a lot, and there is a sudden increase compared to weekdays.
In terms of usage time, customers are clearly higher than subscribers.

### The relationship between weekdays and others in 2020

-   Weekdays vs number of trip

```{r}
library(ggplot2)
ggplot(trip_2020, aes(x = day_of_week, y = num_trip_of_weekday, fill = member_casual)) +
  geom_col() + 
  labs(title = "The number of weekday 2020")
```

-   Weekdays vs trip time

```{r}
library(ggplot2)
ggplot(trip_2020, aes(x = day_of_week, y = mean_time_trip, fill = member_casual)) +
  geom_col() + 
  labs(title = "The trip time of weekday 2020")
```

*Similar to 2019, in 2020 the data still tells the same story, the number of uses of the member group is mainly during the week, and on weekends there is a clear increase from casual people.* In terms of distance traveled, casual people travel longer, longer than members (move quickly, at a time).
=\> the casual group tends to relax and entertain.
The member group tends to go to work, move quickly.

### The starting time

Creating the new column dimentrating the starting time in each day of user.

```{r}
library(lubridate)
divvy_2019_q1 <- divvy_2019_q1 %>%
  mutate(start_hour = hour(start_time))
divvy_2020_q1 <- divvy_2020_q1 %>% 
  mutate(start_hour = hour(started_at))
```

What the relationship of the start time in day with type of user?
\* 2019:

```{r}
ggplot(divvy_2019_q1, aes( x = start_hour, fill = usertype)) +
  geom_bar()
```

-   2020:

```{r}
ggplot(divvy_2020_q1, aes( x = start_hour, fill = member_casual)) +
  geom_bar()
```

In the both graph, the same trends that casual (customer -2019) use the firm service in the 10-16h.
The member (subcriber - 2019) user service in the time to going work almost

## Conclusion

1.  Launch Targeted Weekend Membership Promotions Insight: Casual riders mostly ride on weekends and have longer ride durations, indicating leisure-oriented behavior.
    Action: Develop a “Weekend Explorer Membership” — a low-cost, short-term plan (e.g., valid only on Saturdays and Sundays) promoted via mobile notifications and email after a casual ride on a weekend.
    This taps directly into their behavior and encourages habitual use.

2.  Trigger Smart Notifications After Ride Thresholds Insight: Casual riders already interact with the system multiple times before considering membership.
    Action: Implement automated in-app messages or emails after the 2nd or 3rd casual ride, highlighting the cost savings and benefits of becoming a member.
    Include a comparison: “You’ve already spent \$X this week — a membership could’ve saved you Y%.”

3.  Create Personalized Onboarding for Casual Riders Insight: Casual users are diverse — some explore the city, others exercise.
    Action: After a casual rider’s first trip, ask one quick question via app or email: “What do you use our bikes for?”
    Based on the response (e.g., leisure, fitness, commuting), deliver custom benefits messaging tied to their habits — e.g., fitness perks, commuter pass incentives, or scenic route tips — all while nudging toward membership.
